---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r setup}

  library(envReport)
  library(envImport)
  library(magrittr)

```

# envImport

<!-- badges: start -->
<!-- badges: end -->

The goal of envImport is to obtain, and assemble, environmental data from disparate data sources, for a geographic area of interest. As little filtering as possible occurs when obtaining data and envImport does _not_ aim to clean, filter or tidy the data (see [envClean](https://acanthiza.github.io/envClean/) for help there). Usually the end result of an envImport workflow is a single object that contains all records from all the data sources. Sourcing and assembling environmental rasters for the area of interest is also in scope, but poorly implemented / documented currently.

## Installation

You can install the development version of envImport from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Acanthiza/envImport")
```

## Supported data sources

`data_name` = 'data source'. Data sources are (usually) obvious sources of data. Examples are the Global Biodiversity Infrastructure Facility ([GBIF](https://www.gbif.org/)) or Terrestrial Ecosystems Network ([TERN](https://www.tern.org.au/)). The `r nrow(envImport::data_map)` data sources currently supported or with plans for development are (also see `envImport::data_map`):

* `r paste0(data_map$data_name, ": ", data_map$desc, collapse = "\n* ")`

## General workflow

### `data_map`

The data_map (see table @ref(tab:dataMap)) provides a mapping from original data sources to the desired columns in the assembled data set.

```{r dataMap}

  knitr::kable(data_map
               , caption = "Data map of desired columns in the assembled data (columns) and names of columns in the original data (rows)"
               )


```

### `get_x`

`get_x` functions get data from the data source `x`. Results are always saved to disk (as getting data can be slow). When run again, they load from the saved file by default. If available, `get_x` functions use any R packages and functions provided by the data source (e.g. GBIF provides `rgbif` `r cite_package("rgbif")` and TERN provides `ausplotsR` `r cite_package("ausplotsR")`). The first arguments to `get_x` functions are always:

* `aoi`: an area of interest, provided as simple feature. `get_x` will turn any `aoi` into a bounding box and convert to coordinates appropriate for data source `x`. [Ed: is `aoi` always required?]
* `save_dir`: a directory to save the results to. The default (`NULL`) leads to the file `here::here("out", "ds", "x.rds")` being created and used as `save_file`. `ds` is for 'data source'. While the saved file is usually `x.rds`, in some instances it follows the format and naming of the download from `x` (e.g. GBIF data comes in a `.zip` file named by the corresponding download key).
* `get_new`: an override to force `get_x` to requery the data source, even if save_file already exists
* `...`: the dots are passed to any underlying 'native' function, such as `rgbif::occ_download()` or `ausplotsR::get_ausplots()`

Many of the `get_x` functions will only work within `DEW`.

### `make_x`

`make_x` functions take the results from `get_x` functions and prepare them for assembly. This usually involves:

* selecting columns (this is really just aesthetic and helps with debugging when there are tens of columns returned)
* munging a column or two to align with the data_map. For example, the raw TERN plant data obtained via get_tern is point intercept data. The `make_tern` function transforms the raw data into projected foliage cover data (making use of the `ausplotsR::species_table()` function)

### Assemble


